Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance

### 2023-01-04 text/Manuscript.Rmd
edit a part of manuscripts

![**Figure 1. Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** (A) SEM image of a 3-day-old Platynereis larva. (B and C) Five types of neurons (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a 3-day-old larva. The INNOS on the ventral side labelled with (C) are shown in purple. Axons and dendrites appear as lines and cell body positions are represented by spheres. (D) Wiring diagram of the cPRC circuit. (E-G) Expression analysis of the NOS gene (magenta) using in situ HCR. (E) Dorsal view of the larva at 3-day-old. (F and G) Anterior view of the larva at 3 (F) and 2-day-old (G). (F) Insert: showing two NOS cells on each side. (G) Co-stained image using acetylated α-tubulin antibody (acTub: green). (H) Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (NOSp::palmi-3xHA: magenta). Co-staining with acetylated α-tubulin antibody as marker for cilia of cPRC and axonal scaffold (acTub: green). Views of larvae at 3-day-old from the anterior side.](../figures/Fig1.png){width="100%"}

@@

![**Figure 2. Larvae with inhibited NOS have shorter ciliary arrest times and a delayed response to UV-avoidance.** (A) Vertical swimming in larvae treated with NOS inhibitors, L-NAME at 3-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top. The data are shown in 30 s bins. (B) Top: The domain and Exon/Intron structure of NOS. Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9. The generated mutants (NOSΔ11, NOSΔ23) are also shown. Pink indicates target sites. Gray shows PAM sequences, red shows stop codons. (C and D) Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at 2-day-old (C) and 3-day-old (D) stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top. The data are shown in 30 s bins.

(E and F) Overlaid trajectories for WT (left, n=20) and NOS mutant (right, n=31) 3-day-old larvae. Grey shows the trajectories before UV stimulation and purple shows the trajectories after UV stimulation.

(E) The temporal changes in the vertical position of the WT and mutant larvae before and after UV stimulation are shown. The starting points of each larval trajectory are set to 0.

(F) The temporal changes in the distance traveled of the WT and mutant larvae before and after UV stimulation are shown.

(E and F) After UV stimulation is indicated by purple squares.](../figures/Fig2.png){width="100%"}

@@

![**Figure 4. cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** (A) Expression analysis of the NIT-GC1 gene (magenta) using in situ HCR. White arrowheads indicate cells expressing NIT-GC1.Co-staining image using acetylated α-tubulin antibody (ac α-tub: green). (B) Co-expression analysis image of the c-opsin1 gene (green). Black arrowheads indicate cells expressing c-opsin1 differently from cPRC. (C and D) Localisation analysis using NIT-GC1 and NIT-GC2 antibodies. Green shows co-staining with acetylated α-tubulin antibody (ac α-tub).

(E-G) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in control (non-injection) and NIT-GC1 and NIT-GC2 knockdown by morpholino (MO1 and 2) larvae.

(H) The domain structure of NIT-GC1. Grey squares indicate the transmembrane domain.
(I) A schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control. Green cGull shows cGMP-dependent green fluorescence. (J-M) The intensity changes in cGMP fluorescence over time for the four conditions are shown. Gray traces show the optical responses to individual intensities (n \> 6). The individual intensities are normalized (ΔF/F0). The dark line shows the mean value and the band the standard error range of the mean. The respective solution was acted on 2 min after the start of imaging (grey bars).](../figures/Fig4.png){width="100%"}

@@

![**Figure 5. NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** (A) Schematic diagram of the immunostaining procedure after calcium imaging. (B and D) The position of the cPRC in calcium imaging with GCaMP6s. Cells around the cPRC are numbered according to the position of the nuclei. (C and E) Immunostaining image using RYamide (C) or RGWamide and 5-HT antibodies (red) (E) and DAPI (cyan) in the same region as (B) or (D). IN-NOS, IN-RGWa and Ser-h1 identified from the numbered nuclear positions are shown. (F) Correlation (Pearson's r) map of neuronal activity of the cPRC, IN-NOS, IN-RGWa and Ser-h1 neurons.

(E) Changes over time in GCaMP6s intensities in IN-NOS when cPRC is stimulated with UV in WT (dark purple) and NOS mutant (light purple) larvae. (F) Changes over time in GCaMP6s intensities in IN-NOS when cPRC is stimulated with UV in NIT-GC2 morphants larvae. (H) Changes over time in GCaMP6s intensities in IN-RGWa when cPRC is stimulated with UV in WT (dark blue) and NOS mutant (light blue) larvae. (I) Changes over time in GCaMP6s intensities in Ser-h1 when cPRC is stimulated with UV in WT (dark red) and NOS mutant (light red) larvae.](../figures/Fig5.png){width="100%"}

@@

![**Figure 5. Gene expressions and signalling mechanisms between cPRC, INNOS and INRGWa.** (A) Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq. The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells. The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene. (B) Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.](../figures/Fig5.png){width="100%"}

>>>>

![**Figure 1. Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** SEM image of a 3-day-old Platynereis larva. **(B and C)** Five types of neurons (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a 3-day-old larva. Ciliated cells labelled with (B) are shown in grey. The INNOS on the ventral side labelled with (C) are shown in purple. Axons and dendrites appear as lines and cell body positions are represented by spheres. **(D)** Expression analysis of the NOS gene (magenta) using in situ HCR. Anterior view of the larva at 2-day-old. Co-stained image using acetylated α-tubulin antibody (acTub: green). **(E)** Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (NOSp::palmi-3xHA: magenta). Co-staining with acetylated α-tubulin antibody as marker for cilia of cPRC and axonal scaffold (acTub: green). Views of larvae at 3-day-old from the anterior side. **(F)** Wiring diagram of the cPRC circuit. Hexagons and arrows represent neuron groups and their synaptic connections, respectively. Numbers inside hexagons indicate the number of neurons grouped in each node. Interactions with less than five synapses were filtered out in this display. Arrow line thickness (synapse weight) is equal to the common logarithm of the number of synapses (scale is shown on the bottom left of the panel).](../figures/Fig1.png){width="100%"}

@@

![**Figure 2. Larvae with inhibited NOS have shorter ciliary arrest times and a delayed response to UV-avoidance.** (A) Vertical swimming in larvae treated with NOS inhibitors, L-NAME at 3-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top. The data are shown in 30 s bins. (B) Top: The domain and Exon/Intron structure of NOS. Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9. The generated mutants (NOSΔ11, NOSΔ23) are also shown. Pink indicates target sites. Gray shows PAM sequences, red shows stop codons. (C and D) Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at 2-day-old (C) and 3-day-old (D) stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top. The data are shown in 30 s bins. (E and F) Overlaid trajectories for WT (left, n=20) and NOS mutant (right, n=31) 3-day-old larvae. Grey shows the trajectories before UV stimulation and purple shows the trajectories after UV stimulation. (E) The temporal changes in the vertical position of the WT and mutant larvae before and after UV stimulation are shown. The starting points of each larval trajectory are set to 0.(F) The temporal changes in the distance traveled of the WT and mutant larvae before and after UV stimulation are shown.(E and F) After UV stimulation is indicated by purple squares.](../figures/Fig2.png){width="100%"}

@@

![**Figure 4. cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** (A) Expression analysis of the NIT-GC1 gene (magenta) using in situ HCR. White arrowheads indicate cells expressing NIT-GC1.Co-staining image using acetylated α-tubulin antibody (ac α-tub: green). (B) Co-expression analysis image of the c-opsin1 gene (green). Black arrowheads indicate cells expressing c-opsin1 differently from cPRC. (C and D) Localisation analysis using NIT-GC1 and NIT-GC2 antibodies. Green shows co-staining with acetylated α-tubulin antibody (ac α-tub).(E-G) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in control (non-injection) and NIT-GC1 and NIT-GC2 knockdown by morpholino (MO1 and 2) larvae.(H) The domain structure of NIT-GC1. Grey squares indicate the transmembrane domain.(I) A schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control. Green cGull shows cGMP-dependent green fluorescence. (J-M) The intensity changes in cGMP fluorescence over time for the four conditions are shown. Gray traces show the optical responses to individual intensities (n \> 6). The individual intensities are normalized (ΔF/F0). The dark line shows the mean value and the band the standard error range of the mean. The respective solution was acted on 2 min after the start of imaging (grey bars).](../figures/Fig4.png){width="100%"}

@@

![**Figure 5. NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** (A) Schematic diagram of the immunostaining procedure after calcium imaging. (B and D) The position of the cPRC in calcium imaging with GCaMP6s. Cells around the cPRC are numbered according to the position of the nuclei. (C and E) Immunostaining image using RYamide (C) or RGWamide and 5-HT antibodies (red) (E) and DAPI (cyan) in the same region as (B) or (D). IN-NOS, IN-RGWa and Ser-h1 identified from the numbered nuclear positions are shown. (F) Correlation (Pearson's r) map of neuronal activity of the cPRC, IN-NOS, IN-RGWa and Ser-h1 neurons.(E) Changes over time in GCaMP6s intensities in IN-NOS when cPRC is stimulated with UV in WT (dark purple) and NOS mutant (light purple) larvae. (F) Changes over time in GCaMP6s intensities in IN-NOS when cPRC is stimulated with UV in NIT-GC2 morphants larvae. (H) Changes over time in GCaMP6s intensities in IN-RGWa when cPRC is stimulated with UV in WT (dark blue) and NOS mutant (light blue) larvae. (I) Changes over time in GCaMP6s intensities in Ser-h1 when cPRC is stimulated with UV in WT (dark red) and NOS mutant (light red) larvae.](../figures/Fig5.png){width="100%"}

@@

![**Figure 6. Gene expressions and signalling mechanisms between cPRC, INNOS and INRGWa.** (A) Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq. The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells. The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene. (B) Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.](../figures/Fig6.png){width="100%"}
